---
title: "Stats1 Assessment"
author: "Aaron Munro"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Main Findings

## Executive Summary

## Dataset Overview
The dataset I chose to use for the assessment was found on Kaggle. Over a period of a year in 2015, data was collected containing the beer consumption of students around a university area in Sao Paulo, Brazil.

The original columns were in Brazilian so I have translated them for this project:

- Data - Date
- Temperatura Media (C) - Median Temp. (C)
- Temperatura Minima (C) - Min. Temp (C)
- Temperatura Maxima (C) - Max. Temp (C)
- Precipitacao (mm) - Precipitation (mm)
- Final de Semana - Weekend
- Consumo de cerveja (litros) - Consumption (litres)


## Hypothesis
Does an increase in temperature lead to an increase in beer consumption among students in Sao Paulo, Brazil?

H0 - Increased temperatures do not lead to an increase in beer consumption.

H1 - Increased temperatures do lead to an increase in beer consumption

``` {r imports, echo=FALSE}
# Import packages
library(tidyverse)
library(patchwork)
library(lubridate)
library(corrplot)
```

## Original Dataset Analysis

The dataset I chose had a couple of formatting issues I needed to resolve before analysing and simulating the data. First of all, a number of columns had a comma instead of a decimal point. This was easy to fix with a modifier on the ```read.csv()``` function. I also had to remove a large amount of empty rows and finally format each of the columns correctly.

``` {r orig_read, echo=FALSE}
# Import data - dataset has commas instead of decimal points and many blank rows
orig_data <- read.csv("data/Consumo_cerveja.csv", dec = ",")
orig_data <- orig_data[1:365,]

set.seed(7)
```

``` {r orig_rename, echo=FALSE}
# Rename columns
orig_data <- orig_data %>%
  rename(date = Data,
         temp_median = Temperatura.Media..C.,
         temp_min = Temperatura.Minima..C.,
         temp_max = Temperatura.Maxima..C.,
         precip = Precipitacao..mm.,
         weekend = Final.de.Semana,
         consum = Consumo.de.cerveja..litros.)
```

``` {r orig_head, echo=FALSE}
head(orig_data)
```

``` {r orig_comma_rep, echo=FALSE}
# Format fields correctly
orig_data$date <- as_date(as.Date(orig_data$date))
orig_data$temp_median <- as.numeric(orig_data$temp_median)
orig_data$temp_min <- as.numeric(orig_data$temp_min)
orig_data$temp_max <- as.numeric(orig_data$temp_max)
orig_data$precip <- as.numeric(orig_data$precip)
orig_data$weekend <- as.factor(orig_data$weekend)
orig_data$consum <- as.numeric(as.character(orig_data$consum))
```

``` {r orig_summary, echo=FALSE}
summary(orig_data)
```

``` {r orig_add_month, echo=FALSE}
# Add month column
orig_data <- orig_data %>%
  mutate(mo = month(date))
```

``` {r orig_distr, echo=FALSE}
orig_p1 <- ggplot(orig_data, aes(x = temp_median)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Median Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_median)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_median)),col='green',linewidth=0.5)

orig_p2 <- ggplot(orig_data, aes(x = temp_min)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Min. Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_min)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_min)),col='green',linewidth=0.5)

orig_p3 <- ggplot(orig_data, aes(x = temp_max)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Max. Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_max)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_max)),col='green',linewidth=0.5)

orig_p4 <- ggplot(orig_data, aes(x = precip)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Precipitation",
       x = "Precipitation (mm)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(precip)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(precip)),col='green',linewidth=0.5)

orig_p5 <- ggplot(orig_data, aes(x = weekend)) +
  geom_bar() +
  labs(title = "Count of Weekdays vs Weekends",
       x = "Weekend",
       y = "Count")

orig_p6 <- ggplot(orig_data, aes(x = consum)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Beer Consumption",
       x = "Consumption (litres)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(consum)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(consum)),col='green',linewidth=0.5)

orig_patchwork <- orig_p1 + orig_p2 + orig_p3 + orig_p4 + orig_p5 + orig_p6
orig_patchwork + plot_annotation(
  title = 'Original Dataset Overview',
  subtitle = 'Red Line - Mean, Green Line - Median'
)
```

``` {r orig_time, echo=FALSE}
orig_time_p1 <- ggplot(orig_data, aes(x = mo, y = temp_median, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)

orig_time_p2 <- ggplot(orig_data, aes(x = mo, y = temp_min, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)

orig_time_p3 <- ggplot(orig_data, aes(x = mo, y = temp_max, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)

orig_time_p4 <- ggplot(orig_data, aes(x = mo, y = precip, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)

orig_time_p5 <- ggplot(orig_data, aes(x = mo, y = consum, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)

orig_patchwork <- orig_time_p1 + orig_time_p2 + orig_time_p3 + 
  orig_time_p4 + orig_time_p5
orig_patchwork + plot_annotation(
  title = 'Continuous Variables by Month'
)
```

``` {r orig_rain_chance, echo=FALSE}
# Create column for whether it rained on a particular day or not
orig_data <- orig_data %>%
  mutate(rained = ifelse(precip > 0, 1, 0))

# Calculate chance of rain for each month
orig_data %>%
  group_by(mo) %>%
  summarise(rain_days = length(rained[rained==1]),
            mo_days = length(mo),
            rain_chance = length(rained[rained==1])/length(mo))

# Distribution of precipitation on the days it rained
orig_data %>%
  filter(rained == 1) %>%
  ggplot(aes(x = mo, y = precip, group = mo)) +
  geom_boxplot() +
  scale_x_continuous(breaks = seq(1,12,1)) +
  stat_summary(fun = "mean", aes(ymax = after_stat(y), ymin = after_stat(y)), 
               color = "red", shape = 4)
```

## Simulated Data

As we're working with weather data across a year, simulating a random set of data based on the overall distribution of each column would not make sense. I wanted the data to represent the changing of temperature across the year in a more realistic way.

To do this I would gather the relevant statistics (mean, standard deviation) across each month and then create a distribution for the number of days in the month.

### Date Range

The first task is to create the range of dates for 2015 using a package called lubridate.
``` {r sim_date, echo=FALSE}
# Start date
start_date <- ymd("2015/01/01")
  
# End date
end_date <- ymd("2015/12/31")
  
# Create range of dates
dates <- seq(start_date, end_date,"days")
```

### Temperature Columns

For each of the temperature columns I will create a normal distribution for each month then combine them into a vector cotaining a full year of data.

``` {r sim_temp_median, echo=FALSE}
# Get mean and standard deviation for each month
temp_med_stat <- orig_data %>%
  group_by(mo) %>%
  summarise(mean = mean(temp_median), sd = sd(temp_median))

# Create distributions
temp_med_jan <- rnorm(31, mean=temp_med_stat$mean[[1]], sd=temp_med_stat$sd[[1]])
temp_med_feb <- rnorm(28, mean=temp_med_stat$mean[[2]], sd=temp_med_stat$sd[[2]])
temp_med_mar <- rnorm(31, mean=temp_med_stat$mean[[3]], sd=temp_med_stat$sd[[3]])
temp_med_apr <- rnorm(30, mean=temp_med_stat$mean[[4]], sd=temp_med_stat$sd[[4]])
temp_med_may <- rnorm(31, mean=temp_med_stat$mean[[5]], sd=temp_med_stat$sd[[5]])
temp_med_jun <- rnorm(30, mean=temp_med_stat$mean[[6]], sd=temp_med_stat$sd[[6]])
temp_med_jul <- rnorm(31, mean=temp_med_stat$mean[[7]], sd=temp_med_stat$sd[[7]])
temp_med_aug <- rnorm(31, mean=temp_med_stat$mean[[8]], sd=temp_med_stat$sd[[8]])
temp_med_sep <- rnorm(30, mean=temp_med_stat$mean[[9]], sd=temp_med_stat$sd[[9]])
temp_med_oct <- rnorm(31, mean=temp_med_stat$mean[[10]], sd=temp_med_stat$sd[[10]])
temp_med_nov <- rnorm(30, mean=temp_med_stat$mean[[11]], sd=temp_med_stat$sd[[11]])
temp_med_dec <- rnorm(31, mean=temp_med_stat$mean[[12]], sd=temp_med_stat$sd[[12]])

# Combine into a single vector
temp_med <- c(temp_med_jan, temp_med_feb, temp_med_mar, temp_med_apr,
                   temp_med_may, temp_med_jun, temp_med_jul, temp_med_aug,
                   temp_med_sep, temp_med_oct, temp_med_nov, temp_med_dec)
```

``` {r sim_temp_min, echo=FALSE}
# Get mean and standard deviation for each month
temp_min_stat <- orig_data %>%
  group_by(mo) %>%
  summarise(mean = mean(temp_min), sd = sd(temp_min))

# Create distributions
temp_min_jan <- rnorm(31, mean=temp_min_stat$mean[[1]], sd=temp_min_stat$sd[[1]])
temp_min_feb <- rnorm(28, mean=temp_min_stat$mean[[2]], sd=temp_min_stat$sd[[2]])
temp_min_mar <- rnorm(31, mean=temp_min_stat$mean[[3]], sd=temp_min_stat$sd[[3]])
temp_min_apr <- rnorm(30, mean=temp_min_stat$mean[[4]], sd=temp_min_stat$sd[[4]])
temp_min_may <- rnorm(31, mean=temp_min_stat$mean[[5]], sd=temp_min_stat$sd[[5]])
temp_min_jun <- rnorm(30, mean=temp_min_stat$mean[[6]], sd=temp_min_stat$sd[[6]])
temp_min_jul <- rnorm(31, mean=temp_min_stat$mean[[7]], sd=temp_min_stat$sd[[7]])
temp_min_aug <- rnorm(31, mean=temp_min_stat$mean[[8]], sd=temp_min_stat$sd[[8]])
temp_min_sep <- rnorm(30, mean=temp_min_stat$mean[[9]], sd=temp_min_stat$sd[[9]])
temp_min_oct <- rnorm(31, mean=temp_min_stat$mean[[10]], sd=temp_min_stat$sd[[10]])
temp_min_nov <- rnorm(30, mean=temp_min_stat$mean[[11]], sd=temp_min_stat$sd[[11]])
temp_min_dec <- rnorm(31, mean=temp_min_stat$mean[[12]], sd=temp_min_stat$sd[[12]])

# Combine into a single vector
temp_min_df <- c(temp_min_jan, temp_min_feb, temp_min_mar, temp_min_apr,
                   temp_min_may, temp_min_jun, temp_min_jul, temp_min_aug,
                   temp_min_sep, temp_min_oct, temp_min_nov, temp_min_dec)
```

``` {r sim_temp_max, echo=FALSE}
# Get mean and standard deviation for each month
temp_max_stat <- orig_data %>%
  group_by(mo) %>%
  summarise(mean = mean(temp_max), sd = sd(temp_max))

# Create distributions
temp_max_jan <- rnorm(31, mean=temp_max_stat$mean[[1]], sd=temp_max_stat$sd[[1]])
temp_max_feb <- rnorm(28, mean=temp_max_stat$mean[[2]], sd=temp_max_stat$sd[[2]])
temp_max_mar <- rnorm(31, mean=temp_max_stat$mean[[3]], sd=temp_max_stat$sd[[3]])
temp_max_apr <- rnorm(30, mean=temp_max_stat$mean[[4]], sd=temp_max_stat$sd[[4]])
temp_max_may <- rnorm(31, mean=temp_max_stat$mean[[5]], sd=temp_max_stat$sd[[5]])
temp_max_jun <- rnorm(30, mean=temp_max_stat$mean[[6]], sd=temp_max_stat$sd[[6]])
temp_max_jul <- rnorm(31, mean=temp_max_stat$mean[[7]], sd=temp_max_stat$sd[[7]])
temp_max_aug <- rnorm(31, mean=temp_max_stat$mean[[8]], sd=temp_max_stat$sd[[8]])
temp_max_sep <- rnorm(30, mean=temp_max_stat$mean[[9]], sd=temp_max_stat$sd[[9]])
temp_max_oct <- rnorm(31, mean=temp_max_stat$mean[[10]], sd=temp_max_stat$sd[[10]])
temp_max_nov <- rnorm(30, mean=temp_max_stat$mean[[11]], sd=temp_max_stat$sd[[11]])
temp_max_dec <- rnorm(31, mean=temp_max_stat$mean[[12]], sd=temp_max_stat$sd[[12]])

# Combine into a single vector
temp_max_df <- c(temp_max_jan, temp_max_feb, temp_max_mar, temp_max_apr,
                   temp_max_may, temp_max_jun, temp_max_jul, temp_max_aug,
                   temp_max_sep, temp_max_oct, temp_max_nov, temp_max_dec)
```

### Precipitation

``` {r sim_precip, echo=FALSE}
# Get mean and standard deviation for each month
precip_stat <- orig_data %>%
  group_by(mo) %>%
  summarise(mean = mean(precip), sd = sd(precip))

# Create distributions
precip_jan <- rnorm(31, mean=precip_stat$mean[[1]], sd=precip_stat$sd[[1]])
precip_feb <- rnorm(28, mean=precip_stat$mean[[2]], sd=precip_stat$sd[[2]])
precip_mar <- rnorm(31, mean=precip_stat$mean[[3]], sd=precip_stat$sd[[3]])
precip_apr <- rnorm(30, mean=precip_stat$mean[[4]], sd=precip_stat$sd[[4]])
precip_may <- rnorm(31, mean=precip_stat$mean[[5]], sd=precip_stat$sd[[5]])
precip_jun <- rnorm(30, mean=precip_stat$mean[[6]], sd=precip_stat$sd[[6]])
precip_jul <- rnorm(31, mean=precip_stat$mean[[7]], sd=precip_stat$sd[[7]])
precip_aug <- rnorm(31, mean=precip_stat$mean[[8]], sd=precip_stat$sd[[8]])
precip_sep <- rnorm(30, mean=precip_stat$mean[[9]], sd=precip_stat$sd[[9]])
precip_oct <- rnorm(31, mean=precip_stat$mean[[10]], sd=precip_stat$sd[[10]])
precip_nov <- rnorm(30, mean=precip_stat$mean[[11]], sd=precip_stat$sd[[11]])
precip_dec <- rnorm(31, mean=precip_stat$mean[[12]], sd=precip_stat$sd[[12]])

precip_df <- c(precip_jan, precip_feb, precip_mar, precip_apr,
                   precip_may, precip_jun, precip_jul, precip_aug,
                   precip_sep, precip_oct, precip_nov, precip_dec)

# Combine into a single vector
precip_df <- replace(precip_df, precip_df < 0, 0)
```

``` {r sim_consum, echo=FALSE}
# Get mean and standard deviation for each month
consum_stat <- orig_data %>%
  group_by(mo) %>%
  summarise(mean = mean(consum), sd = sd(consum))

# Create distributions
consum_jan <- rnorm(31, mean=consum_stat$mean[[1]], sd=consum_stat$sd[[1]])
consum_feb <- rnorm(28, mean=consum_stat$mean[[2]], sd=consum_stat$sd[[2]])
consum_mar <- rnorm(31, mean=consum_stat$mean[[3]], sd=consum_stat$sd[[3]])
consum_apr <- rnorm(30, mean=consum_stat$mean[[4]], sd=consum_stat$sd[[4]])
consum_may <- rnorm(31, mean=consum_stat$mean[[5]], sd=consum_stat$sd[[5]])
consum_jun <- rnorm(30, mean=consum_stat$mean[[6]], sd=consum_stat$sd[[6]])
consum_jul <- rnorm(31, mean=consum_stat$mean[[7]], sd=consum_stat$sd[[7]])
consum_aug <- rnorm(31, mean=consum_stat$mean[[8]], sd=consum_stat$sd[[8]])
consum_sep <- rnorm(30, mean=consum_stat$mean[[9]], sd=consum_stat$sd[[9]])
consum_oct <- rnorm(31, mean=consum_stat$mean[[10]], sd=consum_stat$sd[[10]])
consum_nov <- rnorm(30, mean=consum_stat$mean[[11]], sd=consum_stat$sd[[11]])
consum_dec <- rnorm(31, mean=consum_stat$mean[[12]], sd=consum_stat$sd[[12]])

# Combine into a single vector
consum_df <- c(consum_jan, consum_feb, consum_mar, consum_apr,
                   consum_may, consum_jun, consum_jul, consum_aug,
                   consum_sep, consum_oct, consum_nov, consum_dec)
```

``` {r sim_data, echo=FALSE}
# Create the data
sim_data <- tibble(
  date = dates,
  temp_median = temp_med,
  temp_min = temp_min_df,
  temp_max = temp_max_df,
  precip = precip_df,
  consum = consum_df
) %>%
  mutate(across(where(is.numeric), round, digits = 2))

# Add month column
sim_data <- sim_data %>%
  mutate(mo = month(date))

# Add the weekend column
sim_data <- sim_data %>%
  mutate(weekend = ifelse(wday(date, week_start = 1) %in% c(6, 7), 1, 0))
```

``` {r sim_distr, echo=FALSE}
sim_dist_p1 <- ggplot(sim_data, aes(x = temp_median)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Median Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_median)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_median)),col='green',linewidth=0.5)

sim_dist_p2 <- ggplot(sim_data, aes(x = temp_min)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Min. Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_min)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_min)),col='green',linewidth=0.5)

sim_dist_p3 <- ggplot(sim_data, aes(x = temp_max)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Max. Temperature",
       x = "Temperature (C)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(temp_max)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(temp_max)),col='green',linewidth=0.5)

sim_dist_p4 <- ggplot(sim_data, aes(x = precip)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Precipitation",
       x = "Precipitation (mm)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(precip)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(precip)),col='green',linewidth=0.5)

sim_dist_p5 <- ggplot(orig_data, aes(x = weekend)) +
  geom_bar() +
  labs(title = "Count of Weekdays vs Weekends",
       x = "Weekend",
       y = "Count")

sim_dist_p6 <- ggplot(sim_data, aes(x = consum)) +
  geom_histogram(bins = 10) +
  labs(title = "Distribution of Beer Consumption",
       x = "Consumption (litres)",
       y = "Count") +
  geom_vline(aes(xintercept = mean(consum)),col='red',linewidth=0.5) +
  geom_vline(aes(xintercept = median(consum)),col='green',linewidth=0.5)

sim_patchwork_1 <- sim_dist_p1 + sim_dist_p2 + sim_dist_p3 + sim_dist_p4 +
  sim_dist_p5 + sim_dist_p6
sim_patchwork_1 + plot_annotation(
  title = 'Simulated Dataset Overview',
  subtitle = 'Red Line - Mean, Green Line - Median'
)
```

## Analysis of Simulated Data


### Correlations

``` {r sim_corr_plot, echo=FALSE}
# Select the columns
cor_matrix <- sim_data %>%
  select(temp_median:weekend)

# Create the correlation matrix
cor_matrix <- round(cor(cor_matrix, use = "complete.obs"), 2)

# Show matrix
cor_matrix

# Plot matrix
corrplot(cor_matrix, method="circle", type = "upper")
```

## Conclusion


## References



