---
title: "Stats1 Assessment"
author: "Aaron Munro"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Main Findings

## Executive Summary

## Dataset Overview
The dataset I chose to use for the assessment was found on Kaggle. Over a period of a year in 2015, data was collected that contains the details of beer consumption of students around a university area in Sao Paulo, Brazil.

The original columns were in Brazilian so I have translated them for this project:

- Data - Date
- Temperatura Media (C) - Median Temp. (C)
- Temperatura Minima (C) - Min. Temp (C)
- Temperatura Maxima (C) - Max. Temp (C)
- Precipitacao (mm) - Precipitation (mm)
- Final de Semana - Weekend
- Consumo de cerveja (litros) - Consumption (litres)


## Hypothesis
Does an increase in temperature lead to an increase in beer consumption among students in Sao Paulo, Brazil?

H0 - Increased temperatures do not lead to an increase in beer consumption.

H1 - Increased temperatures do lead to an increase in beer consumption

## R Packages
``` {r imports}
library(tidyverse)
library(patchwork)
```

## Original Dataset Analysis

``` {r orig_read}
orig_data <- read.csv("data/Consumo_cerveja.csv")
```

``` {r orig_rename}
orig_data <- orig_data %>%
  rename(date = Data,
         temp_median = Temperatura.Media..C.,
         temp_min = Temperatura.Minima..C.,
         temp_max = Temperatura.Maxima..C.,
         precip = Precipitacao..mm.,
         weekend = Final.de.Semana,
         consum = Consumo.de.cerveja..litros.)
```

``` {r orig_head}
head(orig_data)
```

## Conclusion

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
